# Tcomp R package
Data from the 2010 Tourism Forecasting competition
[![Travis-CI Build Status](https://travis-ci.org/ellisp/Tcomp-r-package.svg?branch=master)](https://travis-ci.org/ellisp/Tcomp-r-package)

This project provides an R package of convenient access to the data from the tourism forecasting competition described in [Athanasopoulos, Hyndman, Song and Wu (2011)](http://robjhyndman.com/papers/the-tourism-forecasting-competition/). Athanasopoulos et al was originally published in [International Journal of Forecastsing](http://www.forecasters.org/ijf) (2011) 27(3), 822-84.

Currently this package is only available on GitHub:

```{r echo = FALSE}
opts_chunk$set(comment=NA, fig.width=9)
```


```{r eval = FALSE}
devtools::install_github("ellisp/Tcomp-r-package/pkg")
```

The original data are from various tourism organisations and academics around the world and are no identifiable to their original metadata.  The copies used for this package came from [Professor Hyndman's webpage](http://robjhyndman.com/papers/the-tourism-forecasting-competition/).  Transformation into the format used in this package is done in the `/prep/` folder of this GitHub project; but there's no need for most users to look into that

Usage is straight-forward.  All 1,311 series are elements in a single list called `tourism`.  `tourism` is an object of class `Mcomp`, and each individual series is an element of class `Mdata` (these classes are borrowed from [Hyndman's `Mcomp` R package](https://cran.r-project.org/package=Mcomp) and come with convenient print, plot, subset, etc methods).

```{r}
library(Tcomp)
plot(tourism$M4)
plot(forecast(tourism$Q12$x))
```

A wrapper function `forecast_comp` is provided that fits four models from the `forecast` package and returns the mean absolute scaled error for user-provided testing periods.
```{r}
round(forecast_comp(tourism$M4, tests = list(1, 6, 12, 24, 1:12, 1:24), plot = TRUE), 2)
```

This method will also work with the data from the `Mcomp` package:

```{r}
library(Mcomp)
round(forecast_comp(M3[[2000]], tests = list(1, 3, 6, 1:6), plot = FALSE), 2)
```

An extended example in `/r/extended-example/reproduce.R` aims to show how this can be used efficiently to reproduce some of the results in Athanasopoulos but currently fails to come up with the same results.

